% Student Conduct Poliicies
% This file needs to be included in a main LaTeX file

% Policy updated - 1/24/2017

\chapter{Student Conduct and Requirements}

All students enrolled at Iowa State University and are currently in good standing with ISU may join a M:2:I project or propose a project.  All students are invited and encouraged to be involved in Make to Innovate.  The program is managed by the Aerospace Engineering Department through the Make to Innovate Program Coordinator.  All policies and rules are therefore managed and enforced by the Make to Innovate Program Coordinator and the department.  Any questions or concerns with these rules should be made to the Make to Innovate Program Coordinator.  

\section{Student Code of Conduct}
Students in the M:2:I program represent M:2:I, the Aerospace Engineering Department, the College of Engineering and Iowa State University.  It is for this reason we expect students to adhere to a \index{code of conduct} both on campus and off campus.  Students that fail to adhere to these rules may receive a lower grade, a failing grade, or may be removed from the program.  In addition, these policies do not replace the Iowa State University Student Disciplinary Regulations, and some actions may fall under these regulations and associated disciplinary actions.  Any such action that falls under those regulations will be reported to the \index{Dean of Students Office} (DSO) and may have additional disciplinary action taken including and up to expulsion.  Students should read the Iowa State University Student Disciplinary Regulations and you can find these at this URL - \url{http://www.policy.iastate.edu/policy/SDR}  

The following are items that all students in Make to Innovate are expected to follow.

\begin{itemize}
\item All students must follow all signs posted in the M:2:I lab or any other M:2:I facility
\item All students must follow all lab monitor or lab technician instructions
\item All students must follow instructions from the M:2:I Program Coordinator or any other M:2:I faculty/staff
\item All students must follow instructions from their faculty adviser
\item All students must treat all fellow students with respect
\item All students must treat visiting members and members of the public with respect
\end{itemize}

Any case of misconduct will be reviewed by the M:2:I Program Coordinator and/or the Aerospace Engineering Department chair.  

\section{\index{Student Requirements}}
In order to be part of M:2:I, students must meet the following requirements.

\begin{itemize}
\item Must be in good standing with the University
\item Students must be enrolled in either AerE 294X or AerE 494X
\item A student must not be on academic probation
\item A student must be full-time status and must be on campus.
\item Students must have permission from the instructor
\end{itemize}

Students must be on campus or be taking the majority of their course load on campus.  Students are not permitted to take AerE 294X or AerE 494X while also on internship or Co-Op.  The Make to Innovate program is designed and can not be taken as a distance education course.

\section{\index{M:2:I Project Requirements}}
Teams in M:2:I must meet the following requirements in order to form a new team and to maintain the team status of ``in good standing'' with M:2:I.  Teams that fail one or more of these requirements may lose funding and/or membership in M:2:I.  It may also affect the teams grade for the semester in which they are enrolled.
\begin{itemize}
\item All teams must have 100\% of its active members enrolled in 294X or 494X.
\item All projects must have an approved charter on file with M:2:I.
\item If a team is requesting funding, an approved budget must be on file with M:2:I.
\item Teams must have a minimum of 3 students, at least one student must be enrolled AerE 494X.
\item Teams must not have more than 8 students including the team leader.
\end{itemize}

M:2:I is a program that is designed to compliment students academic achievements.  Therefore, we want our students to not only have fun in M:2:I but succeed academically as well.  If a student has a GPA lower than 2.5 or a team member is put on academic probation, they may not be active in M:2:I unless they have met with both their academic advisor and the M:2:I Program coordinator prior to signing up for M:2:I.  Both the academic adviser and the M:2:I coordinator must approve a student being involved in M:2:I before they are active in M:2:I.
